---
title: 'Exercise 1: Data import'
author: "Karsten Donnay"
date: "14.06.2021"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    fig_height: 6
    highlight: tango
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options:
  chunk_output_type: console
---



```{r knitr_init, echo=TRUE, cache=FALSE, warning=FALSE}
library(knitr)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
rm(list = ls())
```


# Guardian Data
In this exercise session we work with a dataset of Guardian news coverage gathered through their API and tweets sent by their official Twitter account last year. In later sessions, we will cover how to compile (and combine) equivalent data ourselves, until then you will work with these sample data also for the assignments, including today's first assignment.

# Various ways to import data {.tabset .tabset-fade .tabset-pills}

For the first few illustrations today we work with our "sampledata", a simplified version of these data containing just a few key variables. The data can be found in the folder "data/" of this exercise.

Here we import the same dataset in 3 common fileformats: an R-data file, a comma separated file, and an Microsoft excel sheet. This sample data has 3 columns, and about 2.500 rows containing only the links and sections of articles from the Guardian.

## R Data

Load with **load()**.

```{r,echo=TRUE}
rm(list = ls())
load("data/sampledata.Rda")
head(sampledata)
dim(sampledata)
```

## Comma separated files

Use **read.csv()**.

```{r,echo=TRUE}
sampledata_csv <- read.csv("data/sampledata.csv")
```

Attention: check the dimensions: only 1 column, but the dataset included 4 columns.
```{r,echo=TRUE}
dim(sampledata_csv)
head(sampledata_csv)
```

Inspect it with a text editor of your choice: you will see that values are not separated by commas, but by semicolons. We adjust the parsing from .csv accordingly:

```{r,echo=TRUE}
sampledata_csv <- read.csv("data/sampledata.csv", sep=";")
dim(sampledata_csv)
```

## Excel sheets

Install and use the **readxl** package and use the **read_xlsx()** command.
```{r,echo=TRUE}
sampledata_xls <- readxl::read_xlsx("data/sampledata.xlsx")
```
 
 
# Basic overiew

To get a basic overview of a dataset, we might use **str()**

```{r,echo=TRUE}
str(sampledata)
```

As mentioned above, **dim()** provides us with a basic overview of how many rows and columns are included in the dataset. 

```{r,echo=TRUE}
dim(sampledata)
```


The **table()** command provides us with an easy overview of the distribution of a dichotomous or categorical variable.

```{r,echo=TRUE}
table(sampledata$sectionId)
```

You can use **hist()** to plot a histogram of a numeric variable and get an overview. 

```{r,echo=TRUE}
load("data/combined.Rda")
hist(combined$favorite_count)
```

You can use **plot()** to plot two variables against each other.

```{r,echo=TRUE}
plot(combined$favorite_count, combined$retweet_count)
```

# Rmarkdown

This file is or is produced by a R Markdown file. You will find a detailed introduction to RMarkdown here: https://bookdown.org/yihui/rmarkdown/ and a summary sheet here: https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf.

In our case, RMarkdown files consist of sections that include text in combination with code ``chunks''. Markdown files allow us to combine text with code. Markdown files keep the syntax simple and use comparable syntax to LaTeX and HTML. Often, the same commands that work in LaTeX work with Markdown as well. Basic text formatting is done with the following commands:

 * you can make text italic by putting \* or \_ around it. E.g. \*text\* looks like *text*
 * you can make text bold by putting \*\*  it. E.g. \*\*text\*\* looks like **text**
 * if you want to resemble code you need to put \` around it. E.g. \`codetext\` looks like `codetext`
  * Sections are introduced by using \#.
  * Lists can be done with using a space and either \*, \-, or \+ and then again a space.
  
Additionally, you need the package `knitr` to compile or "knit" a Markdown File to an output format. In RStudio you can then chose if you want a PDF, HTML or Word Document.

Code chunks also allow for options. Code Chunks are introduced by \`\`\` and also closed like this. After the opening, we need to specify which type of programming language we want to insert (`r`) and then we e.g. can specify if we want to `echo` our code (`TRUE`) or whether we want to omit it (`FALSE`). Other options are e.g. `eval`, which indicates whether we want to evaluate the code in the chunk below or whether we just want to skip it (`eval=FALSE`).


# Git and assignment
Next you may set up your own github account and download or clone the github repository accompanying the lecture and this exercise. You will find the assignment of the first exercise in the folder "assignments/assignment1" under the name "assignment1.Rmd" or its HTML and PDF version. You need to complete this assignment by adding the necessary code to the prepared RMarkdown file. Please change the name to "firstname_lastname_assignment1.Rmd" and upload it in the Dropbox section for "Assignment 1" on **OLAT**. This will also be the submission format for the next assignments as well.